On hyperparameter tuning through Lipschitz global optimization

### 2023-10-03 2023/ITMO paper/elsarticle-template-num.tex
first version of the scheme
time table

Next, we consider the performance of the iOpt, Optuna, and Hyperopt frameworks. Table \ref{tab:3} shows the running time of the frameworks for optimizing the hyperparameters of the SVC algorithm.  A single optimization run was performed with a maximum number of 200 search trials. It is obvious that even in the case of a single run iOpt shows better performance on many datasets compared to Optuna and Hyperopt.

@@

\begin{table}[ht]
\caption{Time of hyperparameter tuning for a single run of algorithms.}
\label{tab:3}
\begin{tabular}{lllllll}
\hline
Dataset &
  \begin{tabular}[c]{@{}l@{}}Optuna \\ random\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}Optuna \\ tpe\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}Optuna \\ cmaes\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}Optuna \\ nsgaii\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}Hyperopt\\ tpe\end{tabular} &
  iOpt \\ \hline
Balance                                                         & 10.8   & 9.2    & 9.0    & 10.6   & 10.3   & \textbf{7.7}    \\
\begin{tabular}[c]{@{}l@{}}Bank \\ Marketing\end{tabular}       & 2888.5 & 3406.5 & \textbf{2704.7} & 3078.6 & 3190.3 & 2963.3 \\
Banknote                                                        & 31.9   & \textbf{12.0}   & 17.7   & 22.1   & 21.7   & 18.4   \\
\begin{tabular}[c]{@{}l@{}}Breast \\ Cancer\end{tabular}        & 9.3    & 7.6    & 6.3    & 8.1    & 8.5    & \textbf{6.0}    \\
CarEvaluation                                                   & 50.9   & 63.3   & 51.0   & 53.1   & 69.5   & \textbf{45.2}   \\
CNAE9                                                           & 295.6  & 207.6  & 318.4  & 262.2  & 253.5  & \textbf{144.9}  \\
\begin{tabular}[c]{@{}l@{}}Credit \\ Approval\end{tabular}      & 14.2   & 12.6   & 13.4   & 13.4   & 15.8   & \textbf{10.8}   \\
Digits                                                          & 132.8  & \textbf{57.5}   & 82.1   & 111.6  & 92.4   & 62.5   \\
Ecoli                                                           & 5.5    & 6.3    & 4.0    & 5.4    & 6.4    & \textbf{3.6}    \\
Parkinsons                                                      & 3.5    & 5.0    & 3.6    & 3.8    & 4.4    & \textbf{2.9}    \\
Semeion                                                         & 218.4  & 119.5  & 157.9  & 135.7  & 193.6  & \textbf{114.2}  \\
\begin{tabular}[c]{@{}l@{}}Statlog \\ Segmentation\end{tabular} & 156.1  & \textbf{55.4}   & 87.1   & 159.2  & 91.6   & 66.5   \\
Wilt                                                            & 70.0   & 75.8   & 68.5   & 67.4   & 78.2   & \textbf{60.9}   \\
Zoo                                                             & 3.0    & 3.9    & 3.0    & 3.1    & 4.2    & \textbf{1.8}    \\ \hline
\end{tabular}
\end{table}

>>>>

Next, we consider the performance of the iOpt, Optuna, and Hyperopt frameworks. Table \ref{tab:time} shows the running time of the frameworks for optimizing the hyperparameters.  A single optimization run was performed with a maximum number of 200 search trials. It is obvious that even in the case of a single run iOpt shows better performance on many datasets compared to Optuna and Hyperopt.

@@

\caption{Time of hyperparameter tuning.}
\label{tab:time}
\begin{tabular}{cccccc}
\textbf{Dataset}                & \textbf{iOpt}     & \textbf{Optuna(tpe)}        & \textbf{Optuna(cmaes)} & \textbf{Optuna(nsgaii)}    & \textbf{Hyperopt(tpe)}      \\ \hline
                                & \multicolumn{5}{c}{\textbf{SVM}}                                                                                                    \\ \hline
\textbf{Balance}                & \textbf{00:00:15} & 00:00:19 ± 0.66\%           & 00:00:17 ± 8.71\%      & 00:00:20 ± 3.00\%          & 00:00:21 ± 4.44\%           \\
\textbf{Bank-marketing}         & \textbf{01:31:18} & 02:02:05 ± 4.12\%           & 01:49:42 ± 8.35\%      & 01:51:06 ± 4.62\%          & 01:52:03 ± 8.25\%           \\
\textbf{Banknote}               & 00:00:35          & \textbf{00:00:25 ± 3.60\%}  & 00:00:32 ± 10.86\%     & 00:00:48 ± 5.89\%          & 00:00:39 ± 4.37\%           \\
\textbf{Breast Cancer}          & \textbf{00:00:11} & 00:00:15 ± 1.38\%           & 00:00:14 ± 3.45\%      & 00:00:16 ± 4.59\%          & 00:00:18 ± 3.79\%           \\
\textbf{CarEvaluation}          & \textbf{00:01:21} & 00:01:48 ± 2.82\%           & 00:01:28 ± 6.69\%      & 00:01:34 ± 2.01\%          & 00:01:44 ± 3.10\%           \\
\textbf{CNAE-9}                 & \textbf{00:05:53} & 00:08:24 ± 18.64\%          & 00:08:30 ± 3.92\%      & 00:08:27 ± 2.28\%          & 00:08:22 ± 5.14\%           \\
\textbf{Credit   Approval}      & \textbf{00:00:19} & 00:00:26 ± 16.46\%          & 00:00:21 ± 7.39\%      & 00:00:23 ± 2.38\%          & 00:00:25 ± 4.47\%           \\
\textbf{Digits}                 & 00:02:07          & \textbf{00:02:01 ± 5.04\%}  & 00:02:47 ± 7.83\%      & 00:03:35 ± 7.10\%          & 00:03:09 ± 4.74\%           \\
\textbf{Ecoli}                  & \textbf{00:00:08} & 00:00:13 ± 2.09\%           & 00:00:10 ± 2.28\%      & 00:00:10 ± 1.66\%          & 00:00:14 ± 3.30\%           \\
\textbf{Parkinsons}             & \textbf{00:00:07} & 00:00:11 ± 1.38\%           & 00:00:08 ± 0.96\%      & 00:00:08 ± 1.43\%          & 00:00:12 ± 4.09\%           \\
\textbf{Semeion}                & \textbf{00:04:03} & 00:04:16 ± 5.22\%           & 00:05:17 ± 3.24\%      & 00:05:51 ± 1.73\%          & 00:05:57 ± 3.06\%           \\
\textbf{Statlog   Segmentation} & 00:02:07          & \textbf{00:01:45 ± 4.02\%}  & 00:02:59 ± 5.75\%      & 00:04:12 ± 6.43\%          & 00:03:05 ± 4.06\%           \\
\textbf{Wilt}                   & \textbf{00:01:51} & 00:02:17 ± 6.65\%           & 00:02:01 ± 9.83\%      & 00:02:07 ± 3.28\%          & 00:02:16 ± 4.78\%           \\
\textbf{Zoo}                    & \textbf{00:00:05} & 00:00:10 ± 2.08\%           & 00:00:07 ± 1.73\%      & 00:00:07 ± 2.42\%          & 00:00:10 ± 3.47\%           \\ \hline
                                & \multicolumn{5}{c}{\textbf{XGBoost}}                                                                                                \\ \hline
\textbf{Balance}                & \textbf{00:01:49} & 00:03:38 ± 10.57\%          & 00:03:09 ± 10.94\%     & 00:02:36 ± 6.39\%          & 00:02:55 ± 11.63\%          \\
\textbf{Bank-marketing}         & 01:46:27          & 01:52:51 ± 17.65\%          & 02:06:44 ± 14.76\%     & 01:51:28 ± 7.79\%          & \textbf{01:43:15 ± 13.30\%} \\
\textbf{Banknote}               & \textbf{00:02:02} & 00:03:36 ± 18.51\%          & 00:03:15 ± 10.33\%     & 00:02:35 ± 6.93\%          & 00:02:45 ± 5.20\%           \\
\textbf{Breast Cancer}          & \textbf{00:01:45} & 00:03:25 ± 26.50\%          & 00:02:33 ± 19.95\%     & 00:02:31 ± 13.55\%         & 00:02:26 ± 8.54\%           \\
\textbf{CarEvaluation}          & 00:10:10          & 00:12:05 ± 8.62\%           & 00:10:53 ± 8.07\%      & 00:07:14 ± 6.79\%          & \textbf{00:07:42 ± 12.51\%} \\
\textbf{CNAE-9}                 & 03:47:42          & 04:55:03 ± 26.76\%          & 05:43:27 ± 14.35\%     & 03:55:49 ± 13.82\%         & \textbf{03:42:39 ± 18.30\%} \\
\textbf{Credit   Approval}      & \textbf{00:01:30} & 00:02:27 ± 23.63\%          & 00:01:54 ± 23.08\%     & 00:01:58 ± 6.65\%          & 00:01:44 ± 13.34\%          \\
\textbf{Digits}                 & 00:48:57          & 00:59:09 ± 21.83\%          & 00:54:27 ± 8.20\%      & 00:43:35 ± 9.93\%          & \textbf{00:46:52 ± 8.24\%}  \\
\textbf{Ecoli}                  & \textbf{00:02:40} & 00:04:03 ± 21.05\%          & 00:03:30 ± 22.08\%     & 00:02:48 ± 8.63\%          & 00:03:05 ± 8.97\%           \\
\textbf{Parkinsons}             & \textbf{00:01:09} & 00:02:32 ± 16.25\%          & 00:01:52 ± 12.30\%     & 00:01:34 ± 4.24\%          & 00:01:45 ± 12.30\%          \\
\textbf{Semeion}                & 01:52:08          & 01:19:57 ± 12.21\%          & 01:42:52 ± 14.47\%     & 02:01:16 ± 9.22\%          & \textbf{01:41:40 ± 7.68\%}  \\
\textbf{Statlog   Segmentation} & 00:34:59          & 00:43:59 ± 13.68\%          & 00:38:13 ± 6.14\%      & \textbf{00:25:03 ± 8.20\%} & 00:32:15 ± 8.61\%           \\
\textbf{Wilt}                   & 00:06:44          & 00:10:42 ± 6.89\%           & 00:08:20 ± 7.94\%      & \textbf{00:06:11 ± 8.76\%} & 00:08:09 ± 8.43\%           \\
\textbf{Zoo}                    & \textbf{00:01:47} & 00:02:53 ± 20.08\%          & 00:02:21 ± 16.16\%     & 00:02:05 ± 4.28\%          & 00:02:13 ± 9.09\%           \\ \hline
                                & \multicolumn{5}{c}{\textbf{MLP}}                                                                                                    \\ \hline
\textbf{Balance}                & \textbf{00:09:29} & 00:18:14 ± 61.97\%          & 00:26:45 ± 11.25\%     & 00:20:55 ± 11.15\%         & 00:15:37 ± 10.06\%          \\
\textbf{Bank-marketing}         & 06:10:05          & \textbf{04:42:08 ± 39.48\%} & 05:44:30 ± 15.70\%     & 07:04:15 ± 10.63\%         & 08:11:02 ± 16.22\%          \\
\textbf{Banknote}               & \textbf{02:07:36} & 02:29:16 ± 44.57\%          & 04:43:15 ± 13.11\%     & 03:20:56 ± 8.58\%          & 02:39:20 ± 11.03\%          \\
\textbf{Breast Cancer}          & \textbf{00:43:14} & 02:48:26 ± 10.35\%          & 02:41:29 ± 15.72\%     & 02:46:45 ± 5.13\%          & 02:11:56 ± 29.01\%          \\
\textbf{CarEvaluation}          & \textbf{02:08:22} & 04:21:15 ± 17.08\%          & 03:04:09 ± 12.40\%     & 03:45:18 ± 9.32\%          & 03:55:44 ± 19.04\%          \\
\textbf{CNAE-9}                 & 07:46:15          & \textbf{07:26:02 ± 9.33\%}  & 13:29:53 ± 11.28\%     & 10:57:27 ± 12.39\%         & 08:32:12 ± 7.19\%           \\
\textbf{Credit   Approval}      & \textbf{00:43:32} & 02:41:30 ± 10.89\%          & 02:24:56 ± 15.16\%     & 02:39:29 ± 4.92\%          & 01:57:22 ± 29.27\%          \\
\textbf{Digits}                 & \textbf{02:58:50} & 10:50:47 ± 7.21\%           & 09:30:39 ± 5.32\%      & 10:19:20 ± 6.95\%          & 08:45:23 ± 29.04\%          \\
\textbf{Ecoli}                  & \textbf{00:05:52} & 00:08:21 ± 16.93\%          & 00:07:03 ± 16.70\%     & 00:08:08 ± 7.34\%          & 00:07:01 ± 12.44\%          \\
\textbf{Parkinsons}             & \textbf{00:16:21} & 00:46:32 ± 34.33\%          & 00:49:54 ± 14.73\%     & 00:44:24 ± 5.05\%          & 00:34:48 ± 14.89\%          \\
\textbf{Semeion}                & \textbf{03:38:23} & 11:26:28 ± 25.20\%          & 10:19:44 ± 7.97\%      & 09:01:10 ± 13.98\%         & 07:55:17 ± 20.55\%          \\
\textbf{Statlog   Segmentation} & \textbf{00:48:57} & 19:27:00 ± 15.47\%          & 17:09:16 ± 11.57\%     & 17:02:03 ± 3.17\%          & 11:20:02 ± 3.89\%           \\
\textbf{Wilt}                   & \textbf{01:23:04} & 17:57:58 ± 28.49\%          & 19:23:37 ± 11.57\%     & 22:06:21 ± 9.34\%          & 12:38:43 ± 3.14\%           \\
\textbf{Zoo}                    & 00:01:46          & 00:01:52 ± 22.83\%          & 00:01:58 ± 7.72\%      & \textbf{00:01:40 ± 8.30\%} & 00:02:01 ± 12.96\%          \\ \hline

%\begin{table}[ht]
%\caption{Time of hyperparameter tuning for a single run of algorithms.}
%\label{tab:3}
%\begin{tabular}{lllllll}
%\hline
%Dataset &
%  \begin{tabular}[c]{@{}l@{}}Optuna \\ random\end{tabular} &
%  \begin{tabular}[c]{@{}l@{}}Optuna \\ tpe\end{tabular} &
%  \begin{tabular}[c]{@{}l@{}}Optuna \\ cmaes\end{tabular} &
%  \begin{tabular}[c]{@{}l@{}}Optuna \\ nsgaii\end{tabular} &
%  \begin{tabular}[c]{@{}l@{}}Hyperopt\\ tpe\end{tabular} &
%  iOpt \\ \hline
%Balance                                                         & 10.8   & 9.2    & 9.0    & 10.6   & 10.3   & \textbf{7.7}    \\
%\begin{tabular}[c]{@{}l@{}}Bank \\ Marketing\end{tabular}       & 2888.5 & 3406.5 & \textbf{2704.7} & 3078.6 & 3190.3 & 2963.3 \\
%Banknote                                                        & 31.9   & \textbf{12.0}   & 17.7   & 22.1   & 21.7   & 18.4   \\
%\begin{tabular}[c]{@{}l@{}}Breast \\ Cancer\end{tabular}        & 9.3    & 7.6    & 6.3    & 8.1    & 8.5    & \textbf{6.0}    \\
%CarEvaluation                                                   & 50.9   & 63.3   & 51.0   & 53.1   & 69.5   & \textbf{45.2}   \\
%CNAE9                                                           & 295.6  & 207.6  & 318.4  & 262.2  & 253.5  & \textbf{144.9}  \\
%\begin{tabular}[c]{@{}l@{}}Credit \\ Approval\end{tabular}      & 14.2   & 12.6   & 13.4   & 13.4   & 15.8   & \textbf{10.8}   \\
%Digits                                                          & 132.8  & \textbf{57.5}   & 82.1   & 111.6  & 92.4   & 62.5   \\
%Ecoli                                                           & 5.5    & 6.3    & 4.0    & 5.4    & 6.4    & \textbf{3.6}    \\
%Parkinsons                                                      & 3.5    & 5.0    & 3.6    & 3.8    & 4.4    & \textbf{2.9}    \\
%Semeion                                                         & 218.4  & 119.5  & 157.9  & 135.7  & 193.6  & \textbf{114.2}  \\
%\begin{tabular}[c]{@{}l@{}}Statlog \\ Segmentation\end{tabular} & 156.1  & \textbf{55.4}   & 87.1   & 159.2  & 91.6   & 66.5   \\
%Wilt                                                            & 70.0   & 75.8   & 68.5   & 67.4   & 78.2   & \textbf{60.9}   \\
%Zoo                                                             & 3.0    & 3.9    & 3.0    & 3.1    & 4.2    & \textbf{1.8}    \\ \hline
%\end{tabular}
%\end{table}
